---
title: "Notebook: comparing models of Aspect"
author: "Lauren Fonteyn"
output:
  html_document:
    theme: cosmo
---

First import all required libraries:

```{r}
library(vcd)
library(lattice)
library(lme4)
library(effects)
library(qpcR)
library(Hmisc)
```

Next, we read the main data file into memory:

```{r}
asp <- read.csv("Aspect_R_bounding_noamb.csv", header = TRUE)
summary(asp)
```

## Models
The following code was used for the unidimensional model (boundedness not included)

```{r}
asp$lex_asp <- relevel(asp$lex_asp, ref="activity")
model1 <- glmer(type ~ lex_asp + period + lex_asp:period + (1|File), family=binomial, data = asp, control=glmerControl(optimizer="bobyqa"))
summary(model1)
plot(allEffects(model1))
```

For the bidimensional model (*with* temporal boundedness), this code was used:

```{r}
asp$lex_asp_b <- relevel(asp$lex_asp_b, ref="activity")
asp$temp_bound <- relevel(asp$temp_bound, ref="unbound")
model2 <- glmer(type ~ lex_asp_b + temp_bound + period + lex_asp_b:period + temp_bound:period + (1|File), family=binomial, data = asp, control=glmerControl(optimizer="bobyqa"))
summary(model2)
plot(allEffects(model2))
```

## Comparing the models

First, we can conduct an ANOVA to compare the two models

```{r}
anova(model1, model2, test="Chisq")
```

To compute the Akaike weight, we use:

```{r}
akaike.weights(c(AIC(model1), AIC(model2)))
```

The following functions yield the C-statistic for each model:

```{r}
probs1 = 1/(1+exp(-fitted(model1)))
somers2(probs1, as.numeric(asp$type)-1)
probs2 = 1/(1+exp(-fitted(model2)))
somers2(probs2, as.numeric(asp$type)-1)
```

## A third model

Finally, we can also create a third model that only uses temporal boundedness as a predictor:

probs3 = 1/(1+exp(-fitted(model3)))
somers2(probs3, as.numeric(asp$type)-1)
```{r}
asp$temp_bound <- relevel(asp$temp_bound, ref="unbound")
model3 <- glmer(type ~ temp_bound + period + temp_bound:period + (1|File), family=binomial, data = asp, control=glmerControl(optimizer="bobyqa"))
summary(model3)
plot(allEffects(model3))
```

The C-statistic for this model can be calculated as follows:

```{r}
probs3 = 1/(1+exp(-fitted(model3)))
somers2(probs3, as.numeric(asp$type)-1)
```

We can use the same functions again to calculate the akaike weights:

```{r}
akaike.weights(c(AIC(model1), AIC(model2), AIC(model3)))
```

And, finally, the ANOVA:

```{r}
anova(model2, model3, test="Chisq")
```